---
title: "Analysing Apple Watch data end to end"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysing Apple Watch data end to end}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# The analysis steps need a real Apple Health export, so they are shown but not
# executed when the vignette is built.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

```{r setup, eval = TRUE}
library(PhysioDevices)
```

This vignette takes an **Apple Health export** (Health app → profile →
*Export All Health Data* → `export.xml`, plus any ECG `.csv` files) all the way
through to per-modality analysis, using `PhysioDevices` for ingestion and
`PhysioWearable` / `PhysioECG` for the analysis.

## 1. Ingest the export

```{r ingest}
ah <- readAppleHealth("apple_health_export/export.xml", tz = "Asia/Tokyo")
ah
appleHealthTypes(ah)          # what modalities are present, with counts and spans
```

`readAppleHealth()` streams the file, so multi-hundred-MB exports are fine. Every
`<Record>` becomes a row (type, source, unit, start/end, value); `<Workout>`
elements become `ah$workouts`.

## 2. Heart rate and heart-rate variability

The watch already stores HRV as `HeartRateVariabilitySDNN`; heart rate is
`HeartRate`. For deeper HRV from the raw beats, use the single-lead ECG (below).

```{r hrv}
hr  <- appleHealthSeries(ah, "HeartRate")                 # bpm, irregular
hrv <- appleHealthSeries(ah, "HeartRateVariabilitySDNN")  # ms (SDNN)
summary(hr$value_num)
summary(hrv$value_num)

# as a PhysioExperiment (irregular; timestamps in metadata) for the ecosystem
pe_hr <- appleHealthExperiment(ah, "HeartRate")
```

## 3. Blood oxygen (SpO2)

```{r spo2}
ox <- appleHealthSeries(ah, "OxygenSaturation")           # fraction 0-1
PhysioWearable::spo2Metrics(ox$value_num, time = ox$start)
```

`spo2Metrics()` reports mean, nadir, time-below-90% (T90/CT90) and an oxygen
desaturation index (ODI). Note the Apple Watch takes **periodic spot readings**,
not a continuous trace, so ODI/T90 are approximate unless you feed continuous
overnight oximetry.

## 4. Sleep

The watch scores sleep **stages** itself; `summarizeAppleSleep()` turns them into
per-night clinical metrics (TST, efficiency, WASO, latency, stage minutes).

```{r sleep}
stages <- appleHealthSeries(ah, "SleepAnalysis")          # AsleepCore/Deep/REM, Awake, InBed
PhysioWearable::summarizeAppleSleep(stages)
```

If you instead have raw activity counts (from another wearable), score sleep/wake
with the Cole-Kripke algorithm and summarise it:

```{r sleep-actigraphy}
sw <- PhysioWearable::coleKripke(counts)                  # counts = per-minute activity
PhysioWearable::summarizeSleep(sw)
```

## 5. Activity

```{r activity}
steps  <- appleHealthSeries(ah, "StepCount")
energy <- appleHealthSeries(ah, "ActiveEnergyBurned")
tapply(steps$value_num, as.Date(steps$start), sum)        # steps per day
ah$workouts                                               # per-workout summaries
```

For epoch-level free-living metrics (ENMO, intensity gradient, fragmentation,
HAR → ICF), export raw accelerometry from a Core-Motion app and use
`PhysioWearable::readAccelCSV()` → `summarizeFreeLiving()` (Apple Health itself
stores only *derived* activity, not raw acceleration).

## 6. ECG

Apple Watch ECG recordings export as one CSV each; `readAppleECG()` reads the
single-lead trace into a `PhysioExperiment` that feeds `PhysioECG`:

```{r ecg}
ecg <- readAppleECG("ecg_2023-05-01.csv")
samplingRate(ecg)                                         # ~512 Hz
# baseline-correct + Pan-Tompkins R-peaks, then HRV/rhythm, via PhysioECG
proc <- PhysioECG::ecgProcess(ecg)
PhysioECG::ecgAnalyze(proc)                               # time/freq/geometric HRV + rhythm
```

## Scope and limitations

- **Raw accelerometry and ECG voltage are not in `export.xml`** — the ECG is a
  separate CSV (`readAppleECG()`); raw motion needs a Core-Motion capture app.
- **SpO2** is periodic spot data → ODI/T90 are approximate.
- **Sleep stages** come from Apple's own on-device scoring; `coleKripke()` is for
  raw-count actigraphy, not a re-scoring of Apple's stages.
- Timestamps carry the export's timezone offset; set `tz` to your local zone.
