---
title: "Analysing Fitbit and Google Pixel data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysing Fitbit and Google Pixel data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# The steps need a real Fitbit archive, so they are shown but not executed.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

```{r setup, eval = TRUE}
library(PhysioDevices)
```

Fitbit exports your data as an account archive (Settings → *Export Your Account
Archive*) or as the Fitbit slice of **Google Takeout** — a directory of per-day
JSON (and some CSV) files. **Google Pixel Watch stores its health data in Fitbit**,
so the same reader covers Pixel Watch. The analysis then reuses the same
`PhysioWearable` / `PhysioECG` functions as the Apple Watch workflow.

## 1. Ingest the archive

```{r ingest}
fb <- readFitbit("Takeout/Fitbit", tz = "Asia/Tokyo")
fb   # which modalities were found, with row counts and spans
```

`readFitbit()` discovers files by name (`heart_rate-*`, `steps-*`, `sleep-*`,
`resting_heart_rate-*`, SpO2), parses the nested JSON, and skips anything it
cannot read (with a warning). Restrict with `what =` if you only need some
modalities.

## 2. Heart rate

```{r hr}
hr <- fitbitSeries(fb, "heart_rate")     # columns: time, bpm (intraday)
summary(hr$bpm)
fb$resting_heart_rate                    # daily resting HR
```

For HRV, Fitbit exports its own daily RMSSD (a `Heart Rate Variability` file); for
beat-to-beat HRV use a Fitbit Sense/Charge ECG recording with `PhysioECG`.

## 3. Blood oxygen (SpO2)

```{r spo2}
PhysioWearable::spo2Metrics(fb$spo2$value, time = fb$spo2$time)
```

## 4. Sleep

Fitbit scores sleep stages itself (`wake`/`light`/`deep`/`rem`).
`summarizeSleepStages()` turns them into per-night clinical metrics — the same
engine used for Apple Watch, just with Fitbit's stage labels:

```{r sleep}
PhysioWearable::summarizeSleepStages(
  fb$sleep,
  asleep_levels = c("light", "deep", "rem"),
  wake_levels   = "wake",
  stage_cols    = c(light = "light", deep = "deep", rem = "rem"))

fb$sleep_summary          # Fitbit's own per-log efficiency / minutes asleep / time in bed
```

## 5. Activity

```{r activity}
steps <- fitbitSeries(fb, "steps")
tapply(steps$value, as.Date(steps$time), sum)     # steps per day
```

For epoch-level free-living metrics (ENMO, intensity, fragmentation, HAR → ICF)
you need raw accelerometry, which Fitbit does not export — only derived activity.

## Google Fit and Health Connect

Phone-level Google data has two other paths, both reusing the same analysis:

```{r googlefit}
# Google Takeout "Fit": granular heart-rate/step JSON + daily-summary CSVs
gf <- readGoogleFit("Takeout/Fit", tz = "Asia/Tokyo")
gf$heart_rate
gf$daily
```

```{r healthconnect}
# Android Health Connect: a per-record-type CSV export
hc <- readHealthConnect("health_connect_export", tz = "Asia/Tokyo")
PhysioWearable::spo2Metrics(hc$spo2$value, time = hc$spo2$time)
PhysioWearable::summarizeSleepStages(
  hc$sleep, asleep_levels = c("light", "deep", "rem"), wake_levels = "awake",
  stage_cols = c(deep = "deep", rem = "rem"))
```

## Scope and limitations

- Covers the Fitbit **account archive / Google Takeout** JSON layout; exact
  filenames vary slightly by export version (the reader keys off filename
  patterns and skips what it cannot parse).
- The Fitbit **Web API** (OAuth2, live JSON) is a separate path not covered here.
- **Google Pixel Watch** = Fitbit data (`readFitbit()`).
- **Google Fit** (`readGoogleFit()`, Takeout) is being retired by Google in favour
  of **Health Connect** (`readHealthConnect()`, per-type CSV export). Health
  Connect has no single official export file, so column names vary by exporter;
  the reader detects them heuristically.
- No raw acceleration and no continuous SpO2 (spot readings) — same limits as the
  Apple Watch workflow.
